All just Greedy for Money? - A prediction and Analysis of the interplay between Graduate Statistics and Salary Prospects in the STEM Field

### 2024-01-28 doc/DataLiteracyStudentProject/paper.tex
visualise coefficients

@@

\subsection{Model creation}\label{sec:methods:var}
A vector autoregressive model is used to analyze relationships between multiple time series.
The model creates a linear combination for each prediction which can be expressed in a matrix-vector notation.
\begin{equation*}
    \resizebox{\linewidth}{!}{%
        $\begin{bmatrix}
            S_{1,n}\\
            S_{2,n}\\
            \vdots \\
            S_{m,n}
        \end{bmatrix} = 
        \begin{bmatrix}
            C_{1,1} & C_{1,2} & \dots &C_{1,n-1}\\
            C_{2,1} & C_{2,2} &\dots & C_{2,n-1}\\
            \vdots & \vdots  & \ddots &\vdots\\
            C_{m,1} & C_{m,2} &\dots &C_{m,n-1}\\
        \end{bmatrix} \cdot
        \begin{bmatrix}
            S_{1,1} & S_{1,2} & \dots & S_{1,n-1}\\
            S_{2,1} & S_{2,2} & \dots & S_{2,n-1}\\
            \vdots & \vdots  & \ddots &\vdots\\
            S_{m,1} & S_{m,2} & \dots & S_{m,n-1}\\
        \end{bmatrix}+
        \begin{bmatrix}
            \epsilon_{1}\\
            \epsilon_{2}\\
            \vdots\\
            \epsilon_{n}\\
        \end{bmatrix}$%
    }
\end{equation*}
The $\epsilon$ vector approximates an error, which is induced by the model. One main benefit of the VAR model is that it also assumes
that $S_{i,j}$ is also influenced by $S_{k,l}$ $i,k \in \{1,\dots, m\}: i\neq k$ and $j,l \in \{1,\dots, n-1\}: j\neq l$.
A VAR model has one major advantage, over other possible approaches, which is its simplicity. Therefore its results remain interpretable.
This model only uses one hyperparameter which describes the amount of past data that is used to make a prediction. This parameter is called the lag.
The vector autoregressive model heavily depends on the stationarity assumption which ensures that the stochastical process stays the same over time. 
In this context, this is assured by the following three axioms:
\begin{enumerate}[label=\Roman*]
    \item The mean $\mu$ is constant
    \item The standard deviation is constant
    \item There is no seasonality in the data
\end{enumerate}
Since we are only interested in predicting two time steps into the future, which is equivalent to one year,
we can safely assume that our data is stationary.

@@

\section{Results}\label{sec:results}
The hyperparameter optimization is always a trade-off between several optimal values for different parameters. 
Our main goal was to predict the students who enroll in study subjects in the STEM sector when using the salary and the
graduates of the given years as prior knowledge. For this reason, we used a lag of 5, because the overall percentage error
was the lowest for the students.
\begin{figure*}
    \includegraphics{fig/prediction_5.pdf}
    \caption{VAR Model Prediction with hyperparameter p=5}
\end{figure*}
During our analysis, we found out, that the salary has a major influence on our prediction. If our model works correctly, then the amount
of predicted students should go down, because fewer students imply that there is a lack of qualified workers on the job market.
This on the other hand leads to an increase in the salary. Now we manipulated the salary data with a fixed $\alpha\in \mathbb{N}$ to increase
the whole time series by a constant factor. We used this approach because the vector autoregressive model internally uses Mean Squared Error as loss function.
The Mean Squared Error squares the difference between the data points, which leads to a stronger influence of large differences.
This high error leads to a wrong minimization, which gives wrong coefficients.
Another more important point is that we don't want to violate stationarity, which ensures that our model predicts in a good manner.\\
\begin{table}[H]
    \centering
    \begin{tabular}{c|c|c|c}
        & students & salary & graduates\\
        \hline
        students & 1.000000 & -0.115442 &  0.111069\\
        \hline
        salary & -0.115442 & 1.000000 & -0.902321\\
        \hline
        graduates & 0.111069 & -0.902321 &  1.000000\\
    \end{tabular}
    \caption{Correlation Matrix of Var(5) model}
\end{table}

@@

\begin{figure}[H]
    \centering
    \includegraphics{fig/prediction_modified_10p_increase_lags_5.pdf}
    \caption{Increased salary by 10 \% in Var(5) model}
\end{figure}

>>>>

\begin{figure*}
    \includegraphics{fig/prediction_5.pdf}
    \caption{VAR Model Prediction with optimal lag value of 5 on the test data set.}
\end{figure*}

@@

\subsection{Model creation}\label{sec:methods:var}
A vector autoregressive model is used to analyze relationships between multiple time series.
The model creates a linear combination for each prediction which can be expressed in a matrix-vector notation.
\begin{equation*}
    \resizebox{\linewidth}{!}{
        $\begin{bmatrix}
            S_{1,n}\\
            S_{2,n}\\
            \vdots \\
            S_{m,n}
        \end{bmatrix} = 
        \begin{bmatrix}
            C_{1,1} & C_{1,2} & \dots &C_{1,n-1}\\
            C_{2,1} & C_{2,2} &\dots & C_{2,n-1}\\
            \vdots & \vdots  & \ddots &\vdots\\
            C_{m,1} & C_{m,2} &\dots &C_{m,n-1}\\
        \end{bmatrix} \cdot
        \begin{bmatrix}
            S_{1,1} & S_{1,2} & \dots & S_{1,n-1}\\
            S_{2,1} & S_{2,2} & \dots & S_{2,n-1}\\
            \vdots & \vdots  & \ddots &\vdots\\
            S_{m,1} & S_{m,2} & \dots & S_{m,n-1}\\
        \end{bmatrix}+
        \begin{bmatrix}
            \epsilon_{1}\\
            \epsilon_{2}\\
            \vdots\\
            \epsilon_{n}\\
        \end{bmatrix}$%
    }
\end{equation*}
The $\epsilon$ vector approximates an error, which is induced by the model. One main benefit of the VAR model is that it also assumes
that $S_{i,j}$ is also influenced by $S_{k,l}$ $i,k \in \{1,\dots, m\}: i\neq k$ and $j,l \in \{1,\dots, n-1\}: j\neq l$.
A VAR model has one major advantage, over other possible approaches, which is its simplicity. Therefore its results remain interpretable.
This model only uses one hyperparameter which describes the amount of past data that is used to make a prediction. This parameter is called the lag.
The vector autoregressive model heavily depends on the stationarity assumption which ensures that the stochastical process stays the same over time. 
In this context, this is assured by the following three axioms:
\begin{enumerate}[label=\Roman*]
    \item The mean $\mu$ is constant
    \item The standard deviation is constant
    \item There is no seasonality in the data
\end{enumerate}
Since we are only interested in predicting two time steps into the future, which is equivalent to one year,
we can safely assume that our data is stationary.

@@

\section{Results}\label{sec:results}
The hyperparameter optimization is always a trade-off between several optimal values for different parameters. 
Our main goal was to predict the students who enroll in study subjects in the STEM sector when using the salary and the
graduates of the given years as prior knowledge. For this reason, we used a lag of 5, because the overall percentage error
was the lowest for the students.

\begin{figure}
    \includegraphics{fig/params_Students.pdf}
    \caption{Coefficients of VAR model to predict students time-series
    for $t$. $t_1$ to $t_5$ are the values of each time series respectively at the lagged time step.
    $t_1$ is the most recent value.}
\end{figure}
During our analysis, we found out, that the salary has a major influence on our prediction. If our model works correctly, then the amount
of predicted students should go down, because fewer students imply that there is a lack of qualified workers on the job market.
This on the other hand leads to an increase in the salary. Now we manipulated the salary data with a fixed $\alpha\in \mathbb{N}$ to increase
the whole time series by a constant factor. We used this approach because the vector autoregressive model internally uses Mean Squared Error as loss function.
The Mean Squared Error squares the difference between the data points, which leads to a stronger influence of large differences.
This high error leads to a wrong minimization, which gives wrong coefficients.
Another more important point is that we don't want to violate stationarity, which ensures that our model predicts in a good manner.\\
\begin{table}[H]
    \centering
    \begin{tabular}{c|c|c|c}
        & students & salary & graduates\\
        \hline
        students & 1.000000 & -0.115442 &  0.111069\\
        \hline
        salary & -0.115442 & 1.000000 & -0.902321\\
        \hline
        graduates & 0.111069 & -0.902321 &  1.000000\\
    \end{tabular}
    \caption{Correlation Matrix of Var(5) model}
\end{table}

@@

\begin{figure}[H]
    \centering
    \includegraphics{fig/prediction_modified_10p_increase_lags_5.pdf}
    \caption{Increased salary by 10\%}
\end{figure}
